---
title: "Results for Pacific Cod"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Results for Pacific Cod}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, include = FALSE, eval = FALSE}
rmarkdown::render("pacific-cod.Rmd")
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  cache_path = "pacific-cod-cache/",
  fig.path = "pacific-cod-figs-cache/",
  fig.width = 7,
  fig.height = 6
)
```

Using `petrale-sole.Rmd` as a template, which used herring which cut down to the
basics from the `results.Rmd` vignette, to look at Pacific Cod results (for Area
3CD from the 2024 SAR).

Have to work out the reference points explicitly, and define them from the
SAR. Do recruitments first. There are seven model runs which are combined as an
ensemble to give the advice. Need to do the ensembling here, plus exclude the first 1000
MCMC samples from each file.

Set up:
```{r setup, cache = FALSE}
# library(hdiAnalysis)
load_all()
library(dplyr)
```

## Wrangle the model outputs

Want to do recruitments (units are 1000s) and spawning biomass. Need to:

- load samples for each model
- remove first 1000 from each
- combine them
- convert recruitment from thousands to millions
- convert biomass from tonnes to 1000 tonnes
- calculate biomass reference points

Also, biomass can be thought of as spawning biomass, with the assumption of
knife-edged selectivity and maturity.

```{r loaddata}
seven_models <- dir("pacific-cod/pacific-cod-2024")
cod_rec_mcmc <- tibble()
cod_biomass_mcmc <- tibble()

for(i in 1:length(seven_models)){
  # i <- 1  # for testing

  # Spawning biomass
  biomass_this_model <- read.csv(paste0("pacific-cod/pacific-cod-2024/",
                                        seven_models[i],
                                        "/iscam_sbt_mcmc.csv")) %>%
    tibble::as_tibble()

  expect_equal(dim(biomass_this_model),
               c(2000, 68))
  # Remove first 1000 and convert units from tonnes to thousand tonnes
  biomass_this_model <- (biomass_this_model[-(1:1000), ] / 1000) %>%
    as_tibble()

  names(biomass_this_model) <- stringr::str_replace(names(biomass_this_model),
                                                    "sbt1_",
                                                    "")
  cod_biomass_mcmc <- rbind(cod_biomass_mcmc,
                            biomass_this_model)

  # Recruitment
  rec_this_model <- read.csv(paste0("pacific-cod/pacific-cod-2024/",
                                    seven_models[i],
                                    "/iscam_rt_mcmc.csv")) %>%
    tibble::as_tibble()

  expect_equal(dim(rec_this_model),
               c(2000, 65))
  # Remove first 1000 and convert units from thousands to millions:
  rec_this_model <- (rec_this_model[-(1:1000), ] / 1000) %>%
    as_tibble()

  names(rec_this_model) <- stringr::str_replace(names(rec_this_model),
                                                "rt1_",
                                                "")
  cod_rec_mcmc <- rbind(cod_rec_mcmc,
                        rec_this_model)
}

cod_biomass_mcmc
cod_rec_mcmc
```

Reference points are:

- The Upper Stock Reference (USR) is the mean of the biomass estimates from 1956–2004;
- The Lower Reference Point (LRP) is the lowest estimated biomass agreed upon as
  an undesirable state to be avoided. For Area 3CD this is the estimated biomass in 1986.

Scale biomass by the reference points, have to do each separately due to
definitions, and do calculations for each MCMC sample and then summarise.

The USR:
```{r scaleusr}
usr_mcmc <- select(cod_biomass_mcmc,
                   `1956`:`2004`)
biomass_with_usr <- mutate(cod_biomass_mcmc,
                           usr = rowMeans(select(cod_biomass_mcmc,
                                                 `1956`:`2004`)))
biomass_over_usr <- (biomass_with_usr / biomass_with_usr$"usr") %>%
  as_tibble() %>%
  select(-"usr")       # Had checked and they were all 1
```

The LRP:
```{r scalelrp}
biomass_over_lrp <- (cod_biomass_mcmc / cod_biomass_mcmc$`1986`) %>%
  as_tibble()
expect_equal(range(summary(biomass_over_lrp$`1986`)),
             c(1, 1))
```


## Analyse a single vector of values

Let's first do just 2021 for consistency with main hake analysis in
manuscript. Do absolute biomass (as shown in Figure 11 of SAR):
```{r rec1}
cred = 0.95
vec <- pull(cod_biomass_mcmc, "2021")

summary(vec)
length(vec)

res_vec <- create_intervals(vec)
summary_table(res_vec)
```

The ETI and HDI plots (Figure 1 in our manuscript) are:
```{r plot, fig.height = 8, echo = FALSE}
x_lim <- c(0, 35)
y_lim = c(0, 0.18)
y_arrow = 0.16
x_lab = "Spawning biomass (1000 t)"

par(mfrow = c(2, 1))
plot(res_vec,
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.1)
     # main = "Equal-tailed Interval")
plot(res_vec,
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.1)
     # main = "Highest Density Interval")
```

You can see some humps which look almost multimodal (meaning multiple local
peaks), presumably arising from the combining of posteriors from seven models.

So a fair change between the ETI and HDI in this case.

The full results are
```{r fullres}
res_vec$intervals %>% a()       # a() is our shorthand function for as.data.frame()
```
for which definitions are given in the help file `?create_intervals`.

## MCMC samples for multiple years

We now present calculations for  multiple years
of MCMC samples of various quantities.

### Absolute spawning biomass

To create ETIs and HDIs for estimates for each year:
```{r hake_mcmc2}
res_all_years <- create_intervals(cod_biomass_mcmc,
                                  credibility = cred)
```
This gives a list object -- see the rendered `results.html` vignette for
details (such as the intervals for all years are convenienty saved in the single
tibble `res_all_years$intervals_all`).

Plot the full time series like in our manuscript figure:
```{r hake_mcmc7, echo = FALSE}
plot(res_all_years,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 200),
     y_tick_by = 10)
# abline(h = 0.4, col = "red", lty = 2)
# abline(h = 0.8, col = "black", lty = 2)
```

So it pulls down the biomass a bit.

Separate panels:
```{r hake_mcmc7_separate, echo = FALSE, fig.height = 12}
plot(res_all_years,
     type = "separate",
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 200),
     y_tick_by = 10)
# abline(h = 0.4, col = "red", lty = 2)
# abline(h = 0.8, col = "black", lty = 2)
```


### Spawning biomass relative to USR

To create ETIs and HDIs for estimates for each year:
```{r, usr}
res_all_years_rel_usr <- create_intervals(biomass_over_usr,
                                          credibility = cred)
```

TODO check out the warnings.

Plot the full time series like in our manuscript figure:
```{r, plotusr, echo = FALSE}
plot(res_all_years_rel_usr,
     xlab = "Year",
     ylab = "Spawning biomass relative to USR",
     ylim = c(0, 3.25),
     y_tick_by = 0.25,
     add_line = 1,
     add_line_col = "green")
```

Separate panels:
```{r, plotusr2, echo = FALSE, fig.height = 12}
plot(res_all_years_rel_usr,
     type = "separate",
     xlab = "Year",
     ylab = "Spawning biomass relative to USR",
     ylim = c(0, 3.25),
     y_tick_by = 0.25,
     add_line = 1,
     add_line_col = "green")
```

### Spawning biomass relative to LRP

To create ETIs and HDIs for estimates for each year, have to remove 1986 since
all MCMC ratios are 1 (since they are scaled by themselves):
```{r, lrp}
res_all_years_rel_lrp <- create_intervals(select(biomass_over_lrp,
                                                 -"1986"),
                                                 credibility = cred)
```

TODO check out the two warnings. Could be distinct HDIs.

Plot the full time series like in our manuscript figure (just leave out 1986 for
now, could manually add it in) TODO
```{r, plotlpr, echo = FALSE}
plot(res_all_years_rel_lrp,
     xlab = "Year",
     ylab = "Spawning biomass relative to LRP",
     ylim = c(0, 12),
     y_tick_by = 0.5,
     add_line = 1,
     add_line_col = "red")
```
Again look fairly symmetric and so not a huge difference between interval types.

Two panels:
```{r, plotlpr2, echo = FALSE, fig.height = 12}
plot(res_all_years_rel_lrp,
     type = "separate",
     xlab = "Year",
     ylab = "Spawning biomass relative to LRP",
     ylim = c(0, 12),
     y_tick_by = 0.5,
     add_line = 1,
     add_line_col = "red")
```


### Recruitment

To create ETIs and HDIs for estimates for each year:
```{r, recruitment}
res_all_years_rec <- create_intervals(cod_rec_mcmc,
                                      credibility = cred)
```

No warnings.

Plot the full time series like in our manuscript figure:
```{r, plotrec, echo = FALSE}
plot(res_all_years_rec,
     xlab = "Year",
     ylab = "Recruitment (millions of fish)",
     ylim = c(0, 160),
     y_tick_by = 10,
     add_line = 0,
     add_line_col = "grey",
     add_line_lty = 1)
```

Some pulling down of intervals, particularly 2009, 2013, and 2022.Again look fairly symmetric and so not a huge difference between interval types.

Plot in two separate panels to compare (you would only have one in the assessment
document, not the combined plot above)
```{r, plotrec2, echo = FALSE, fig.height = 12}
plot(res_all_years_rec,
     type = "separate",
     xlab = "Year",
     ylab = "Recruitment (millions of fish)",
     ylim = c(0, 160),
     y_tick_by = 10,
     add_line = 0,
     add_line_col = "grey",
     add_line_lty = 1)
```


```{r, knitexit, cache = FALSE}
knitr::knit_exit()
```

HERE - it's started ignoring the knit_exit since I did the new line options for
figures. Made these all echo eq FALSE.

Reproduce in the style of Figure 39 the Petrale Sole stock assessment (excluding 2024):
```{r plotlikeassessment, fig.height = 10, echo = FALSE}
to_plot <- res_all_years$intervals_all %>%
  filter(quantity != 2024)
par(mfrow = c(2,1))
plot(to_plot$quantity,
     to_plot$median,
     type = "o",
     pch = 16,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8),
     main = "Equal-tailed Intervals")
polygon(x = c(to_plot$quantity,
              rev(to_plot$quantity)),
        y = c(to_plot$eti_lower,
              rev(to_plot$eti_upper)),
        col = "lightblue",
        border = NA)
# Redo to be on top
points(to_plot$quantity,
       to_plot$median,
       type = "o",
       pch = 16)
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
add_minor_tickmarks(y_tick_by = 1)

plot(to_plot$quantity,
     to_plot$median,
     type = "o",
     pch = 16,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8),
     main = "Highest Density Intervals")
polygon(x = c(to_plot$quantity,
              rev(to_plot$quantity)),
        y = c(to_plot$hdi_lower,
              rev(to_plot$hdi_upper)),
        col = "lightblue",
        border = NA)
# Redo to be on top
points(to_plot$quantity,
       to_plot$median,
       type = "o",
       pch = 16)
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
add_minor_tickmarks(y_tick_by = 1)
```

Adapted from hake vignette: whether the relative spawning biomass is above or below the management reference points
of 0.4 and 0.8 is key in forming stakeholders' perception of the health of the hake
stock.

Years for which ETI is above 0.4 but HDI is below:
```{r whichyears, eval = FALSE}
dplyr::filter(res_all_years$intervals_all,
              eti_lower > 0.4,
              hdi_lower < 0.4) %>%
  select(quantity,
         eti_lower,
         hdi_lower)
```
Thus, the perception of stock status can depend upon the
definition of credible intervals (ETI versus HDI).


Only one year for which lower end of ETI is above 0.8 but HDI is below (the
stock is increasing fairly quickly around this time):
```{r whichyearsusr, eval = FALSE}
dplyr::filter(res_all_years$intervals_all,
              eti_lower > 0.8,
              hdi_lower < 0.8) %>%
  select(quantity,
         eti_lower,
         hdi_lower)
```

Can see from the time series that upper end of both intervals are always above
0.8.

## 50% credible intervals

How about 50% credible intervals like I did quickly for hake (for which results
were very surpising):
```{r int50, eval = FALSE}
res_all_years_50 <- create_intervals(petrale_mcmc,
                                     credibility = 0.5)
```

```{r int50plot, echo = FALSE, eval = FALSE}
plot(res_all_years_50,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 4))
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
```

## 90% credible intervals

How about 90% credible intervals:
```{r int90, eval = FALSE}
res_all_years_90 <- create_intervals(petrale_mcmc,
                                     credibility = 0.9)
```
