---
title: "Results for Pacific Herring"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Results for Pacific Herring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("results-herring.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  cache_path = "herring-cache/",
  fig.path = "herring-figs-cache/",
  fig.width = 7,
  fig.height = 6
)
```

Using our results.Rmd vignette as a template to analyse Pacific Herring results,
first for the Strait of Georgia. Cutting out the explanatory aspects to just
concentrate on results. Also using hdiAnalysis-vignette-herring-sog.Rmd as a
template for the data wrangling aspect (that was done in the summer before
switching back to `density=FALSE`, hence starting again here and just using some
of that where needed). Only currently intended to work locally (on Andy's
machine), as MCMC values not being pushed.

Set up:
```{r setup}
# library(hdiAnalysis)
load_all()
library(dplyr)
```

## Analyse a single vector of values

This is for Strait of Georgia, and recruitments are age-2 in billions of fish.

The herring Markov chain Monte Carlo (MCMC) recruitment
values are saved as a list object (not saved in the pacakge), each object being
the MCMC results. Which are the 4,000 MCMC values used in the assessment:

DFO (2024) STOCK STATUS UPDATE WITH APPLICATION OF MANAGEMENT PROCEDURES
FOR PACIFIC HERRING (CLUPEA PALLASII) IN BRITISH COLUMBIA: STATUS IN 2023 AND
FORECAST FOR 2024. Science Response.

Recruitment is shown in Figure 13 for the Strait of Georgia, which we'll look at
here first. Columns are years.

```{r loaddata}
# All regions in a list
herring_mcmc <-
  readRDS("herring/herring_mcmc.rds")

regions_all <- c("HG", "PRD", "CC", "SOG", "WCVI")

region <- "SOG"
i_region <- which(regions_all == region)
```

The MCMC samples of herring recruitment for each year are a tibble (choosing one
region here), with each column corresponding to a year and each of the 4,000
rows representing an MCMC sample:

```{r rec}
herring_recruitment_mcmc <- herring_mcmc[[i_region]]$recruitment
sb_mcmc <- herring_mcmc[[i_region]]$spawning_biomass
```


Let's do 2021 for consistency with main hake analysis in manuscript. Remember these are age-2's, and we're doing 90% intervals.

```{r rec1}
cred = 0.90
rec_2021 <- pull(herring_recruitment_mcmc, "2021")

vec <- rec_2021
summary(vec)
length(vec)

res_vec <- create_intervals(vec)
res_vec
```

## Plot the distribution showing the ETI and HDI

The ETI plot (Figure 1A) is:
```{r plot, fig.height = 4}
x_lim <- c(0, 2.5)
y_lim = c(0, 1.8)
x_lab = "Recruitment (billions of age-2 fish)"

plot(res_vec,
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = 1.6,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.1)
```
and the HDI plot is
```{r plot2, fig.height = 4}
plot(res_vec,
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = 1.6,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.1)
```

The main values of interest are simplified in our customised `summary_table()` function:
```{r tab, results = "asis"}
summary_table(res_vec)
```

So not really a huge change between the ETI and HDI in this case, with just a
slight narrowing of the credible interval, because the distribution is fairly
symmetric. Which is somewhat interesting in itself compared to hake (due to the priors?).


The full results are
```{r fullres}
res_vec$intervals %>% a()       # a() is our shorthand function for as.data.frame()
```
for which definitions are given in the help file `?create_intervals`.

## MCMC samples for multiple years

We now present calculations for  multiple years
of MCMC samples of herring recruitment.

To create ETIs and HDIs for estimates of recruitment for each year:
```{r hake_mcmc2}
res_all_years <- create_intervals(herring_recruitment_mcmc,
                                  credibility = cred)
```
This gives a list object, with element `res_all[[i]]` corresponding to column `i`
of the data. Each `res_all[[i]]` element is itself a list, giving the same
results as above for a single vector (and also with the `$name`
element corresponding the column of the input, a year in this example). So the
2021 results from above are:
```{r hake_mcmc3}
res_all_years$res_all[[69]]
```

For convenience, the intervals for all years are saved in a single tibble:
```{r hake_mcmc4}
res_all_years$intervals_all
```
The first column is called `quantity` for generalisability so the code can be
applied to different types of data; in this example it represents the years

To see the main values of interest, namely the ETIs and HDIs and the difference
between their widths:
```{r hake_mcmc5}
summary_table(res_all_years)
```

The sum of the differences over all years between the width of the ETI and width
of the HDI is:
```{r hake_mcmc6}
res_all_years$intervals_all %>%
  dplyr::pull(width_diff) %>%
  sum()
```
so despite apparent small differences, is still over 1.5 billion fish.

To plot the time series:
```{r hake_mcmc7}
plot(res_all_years,
     xlab = "Year",
     ylab = x_lab)
```

TODO double check the ETIs match those shown in the herring stock assessment, then
recomment this and say (values
manually checked, code is below but not run or printed in vignette):
```{r hake_mcmc8, eval = FALSE, echo = FALSE}
res_all_years$intervals_all %>%
  dplyr::select(eti_lower, eti_upper) %>%
  a() %>%
  round(3) * 1000
```

Change do not look as big as for hake, think because there is less contrast
between high and low recruitment year (hake has some very low values), and the
distributions are more symmetric being further away from 0.

### Quickly do the total reduction in uncertainty calculation for each stock

```{r eachstock}
res_all_stocks <- list()
sum_width_diff_all_stocks <- numeric()
for(i in 1:length(regions_all)){
  herring_recruitment_mcmc_this_stock <- herring_mcmc[[i]]$recruitment
  res_all_years_this_stock <- create_intervals(herring_recruitment_mcmc_this_stock,
                                               credibility = cred)

  res_all_stocks[[i]] <- res_all_years_this_stock

  sum_width_diff_all_stocks[i] <- sum(res_all_years_this_stock$intervals_all$width_diff)
}

regions_all
sum_width_diff_all_stocks
sum_width_diff_all_stocks %>% sum()

# Check the SOG values match
expect_equal(sum_width_diff_all_stocks[i_region],
             sum(res_all_years$intervals_all$width_diff))
```

As might be expected, SOG has the highest recruitments and the highest sum of the
differences between ETIs and HDIs. Although if the distributions are more
symmetric because they further away from 0 than for other regions, this might
not necessarily hold. Could have to look at all distributions for
all stocks, beyond current scope. Total difference is in the realm of 'billions
of fish'.

```{r prd, echo = FALSE, eval = FALSE}
res_all_stocks[[which(regions_all == "PRD")]]$intervals_all

plot(res_all_stocks[[which(regions_all == "PRD")]] )
```

Would have to build full analysis for each stock to fully understand. Beyond the
scope right now.

## MCMC samples for multiple years -- relative spawning biomass

Figure 16 of the 2024 herring SAR shows the estimated spawning biomass for 2024
(assuming no fishing). Presumably shown as density kernel. Credible intervals
are shown, clearly demonstrating differences in the probability density at each
end of the ETI (like our Figure 1A for hake in the manuscript).

Let's do all the calculations in one go, then plot just 2024 ETI and HDI for
each region.

```{r, biomassall}
res_all_stocks_biomass <- list()
sum_width_diff_all_stocks_biomass <- numeric()
for(i in 1:length(regions_all)){
  herring_biomass_mcmc_this_stock <- herring_mcmc[[i]]$spawning_biomass
  res_all_years_this_stock_biomass <-
    create_intervals(herring_biomass_mcmc_this_stock,
                     credibility = cred)

  res_all_stocks_biomass[[i]] <- res_all_years_this_stock_biomass

  sum_width_diff_all_stocks_biomass[i] <- sum(res_all_years_this_stock_biomass$intervals_all$width_diff)
}
```

`res_all_stocks_biomass[[1]]$res_all[[74]]` gives the usual class
`intervals_density` results for stock 1 (HG) and quantity 74 which corresponds
to year 2024.

Do the ETI and HDI density plot for each region in turn, the ETI should match
Figure 16 in the SAR. Axes change for each region, but want them consistent
between the two plots, so keep redefining `x_lim` and `y_lim`.

Haida Gwaii:
```{r, hgbiomass, fig.height = 9}
x_lab = "Projected spawning biomass in 2024 (thousands of tons)"
x_lim <- c(0, 40)
y_lim = c(0, 0.23)
x_minor_ticks_by = 1
y_minor_ticks_by = 0.025
y_arrow = 0.2
par(mfrow = c(2,1))

plot(res_all_stocks_biomass[[1]]$res_all[[74]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)

plot(res_all_stocks_biomass[[1]]$res_all[[74]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)
```

Prince Rupert District:
```{r, prdbiomass, fig.height = 9}
x_lim <- c(0, 150)
y_lim = c(0, 0.03)
x_minor_ticks_by = 10
y_minor_ticks_by = 0.001
y_arrow = 0.028

par(mfrow = c(2,1))
plot(res_all_stocks_biomass[[2]]$res_all[[74]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)

plot(res_all_stocks_biomass[[2]]$res_all[[74]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)
```

Central Coast:
```{r, ccbiomass, fig.height = 9}
x_lim <- c(0, 80)
y_lim = c(0, 0.08)
x_minor_ticks_by = 5
y_minor_ticks_by = 0.005
y_arrow = 0.075

par(mfrow = c(2,1))
plot(res_all_stocks_biomass[[3]]$res_all[[74]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)

plot(res_all_stocks_biomass[[3]]$res_all[[74]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)
```

Strait of Georgia:
```{r, sogbiomass, fig.height = 9}
x_lim <- c(0, 300)
y_lim = c(0, 0.018)
x_minor_ticks_by = 50
y_minor_ticks_by = 0.01
y_arrow = 0.017

par(mfrow = c(2,1))
plot(res_all_stocks_biomass[[4]]$res_all[[74]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)

plot(res_all_stocks_biomass[[4]]$res_all[[74]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)
```

West Coast of Vancouver Island:
```{r, wcvibiomass, fig.height = 9}
x_lim <- c(0, 150)
y_lim = c(0, 0.035)
x_minor_ticks_by = 10
y_minor_ticks_by = 0.01
y_arrow = 0.033

par(mfrow = c(2,1))
plot(res_all_stocks_biomass[[5]]$res_all[[74]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)

plot(res_all_stocks_biomass[[5]]$res_all[[74]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = x_minor_ticks_by,
     y_minor_ticks_by = y_minor_ticks_by)
```

Total reduction in estimates of spawning biomass (in 1000s of tons)
over each region over all years (and then total sum of those):
```{r totalsbreduction}
regions_all
sum_width_diff_all_stocks_biomass
sum_width_diff_all_stocks_biomass %>% sum()
```

Might want to replicate hake Figure 2B, which includes the forecasts:
```{r biomassseries, eval = FALSE, echo = FALSE}
plot(res_all_years_2,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
```

<!-- Whether the relative spawning biomass is above or below the management reference point
of 0.4 is key in forming stakeholders' perception of the health of the hake
stock. As seen in the figure, in 2024 and 2025 the ETI does not dip below 0.4 but
the HDI does. Thus, the perception of current and future stock status can depend upon the
definition of credible intervals (ETI versus HDI). The specific years across the
whole time series for which this happens are:
-->

```{r whichyears, echo = FALSE, eval = FALSE}
dplyr::filter(res_all_years_2$intervals_all,
              eti_lower > 0.4,
              hdi_lower < 0.4)
```

Would be interesting to look at relative spawning biomass (like for hake),
compared to the limit reference point of 0.3 B0.
