---
title: "Results for Petrale Sole"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Results for Petrale Sole}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, include = FALSE, eval = FALSE}
rmarkdown::render("results-petrale-sole.Rmd")
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  cache_path = "petrale-sole-cache/",
  fig.path = "petrale-sole-figs-cache/",
  fig.width = 7,
  fig.height = 6
)
```

Using `results-sog-herring.Rmd` as a template, which itself cut down to the
basics from the `results.Rmd` vignette, to do a quick look at Petrale
Sole. Analysing the MCMC results of the ratio of female spawning stock biomass
to female spawning stock biomass at maximum sustainable yield ($B/B_{msy}$), as
shown in Figure 39 of the working paper.

Set up:
```{r setup}
# library(hdiAnalysis)
load_all()
library(dplyr)
```

## Wrangle the model outputs

These are the full relative biomass values for all years.

```{r loaddata}
petrale_mcmc_raw <-
  readRDS("petrale-sole/PetraleBBMSY.rds") %>%
  tibble::as_tibble()
petrale_mcmc_raw$"Year" <- as.numeric(petrale_mcmc_raw$"Year")
petrale_mcmc_raw <- rename(petrale_mcmc_raw,
                           year = Year)
petrale_mcmc_raw
summary(petrale_mcmc_raw)

# Check all years are present:
years <- unique(petrale_mcmc_raw$year)
years
num_years <- length(years)
num_years
range(years)
range(diff(unique(petrale_mcmc_raw$year)))
expect_equal(diff(range(years)) + 1, num_years)

# However, years are not in order:
unique(diff(petrale_mcmc_raw$year))
sum(diff(petrale_mcmc_raw$year) < 0)  # twice
resets <- which(diff(petrale_mcmc_raw$year) < 0)  # twice
resets
# resets gives the last index of each set of mcmc (because it's from a diff, so
# one less than expected). Need to use later
petrale_mcmc_raw[(resets[1] - 2):(resets[1] + 2), ]
petrale_mcmc_raw[(resets[2] - 2):(resets[2] + 2), ]

resets/num_years
nrow(petrale_mcmc_raw)/num_years

num_mcmc <- nrow(petrale_mcmc_raw)/num_years
# Testing:
#petrale_mcmc <- petrale_mcmc_raw[1:2, ]
#petrale_mcmc <- mutate(petrale_mcmc,
#                       sample = 1:2)
```

From Mackenzie: There should be 26,000 samples per year, since we used an
ensemble model with a 50%/25%/25% weighting. To do that we combined the entire
posterior of the medium M model (14,000 samples) and 50% of the posteriors for
the low and high M models (7,000 samples each). The previous data I sent you is
missing some samples. Use the attached data instead, which is 27,995 samples per
year, still missing 5 samples, since it appears BMSY was not estimated for 5
samples. The years start again because it’s the samples from the next model in
the ensemble.

Agrees with the 27,995 calculation above (number of rows divided by number of
year). So just all the samples. Still need to add repeating sample numbers to
then use `pivot_wider`.

Wrangling:
```{r wrangle}
summary(as.factor(petrale_mcmc_raw$year))

# Number of unique MCMC samples for each year, checking each year is the same
expect_equal(num_mcmc,
             unique(summary(as.factor(petrale_mcmc_raw$year))))

expect_equal(nrow(petrale_mcmc_raw),
             num_mcmc * num_years)

#petrale_mcmc <-
#  petrale_mcmc_raw[1:(resets[1]), ]

# Rearrange in order of years, then assign sample numbers to then use pivot_wider:
petrale_mcmc_raw_samples <- arrange(petrale_mcmc_raw,
                                    year) %>%
  mutate(sample = rep(1:num_mcmc,
                      num_years)) %>%
  arrange(sample, year)
petrale_mcmc_raw_samples
petrale_mcmc_raw_samples %>% tail()
```

Then get in desired format, with years as columns, and don't need sample number.
```{r wrangle2}
petrale_mcmc <- tidyr::pivot_wider(petrale_mcmc_raw_samples,
                                   names_from = year,
                                   values_from = Bratio) %>%
  select(-"sample")         # since now just row number
petrale_mcmc
```

## Analyse a single vector of values

Let's first do just 2021 for consistency with main hake analysis in manuscript:
```{r rec1}
cred = 0.95
vec <- pull(petrale_mcmc, "2021")

summary(vec)
length(vec)

res_vec <- create_intervals(vec)
```

The ETI and HDI plots (Figure 1 in our manuscript) are:
```{r plot, fig.height = 8, echo = FALSE}
x_lim <- c(0, 7)
y_lim = c(0, 0.6)
y_arrow = 0.55
x_lab = "Relative biomass"

par(mfrow = c(2, 1))
plot(res_vec,
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.1)
     # main = "Equal-tailed Interval")
plot(res_vec,
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.1)
     # main = "Highest Density Interval")
```

Interesting that you can't tell that the MCMC samples are combined from three
models (there are not obvious humps).

So a fair change between the ETI and HDI in this case.

The main values of interest are simplified in our customised `summary_table()` function:
```{r tab, results = "asis"}
summary_table(res_vec)
```

The full results are
```{r fullres}
res_vec$intervals %>% a()       # a() is our shorthand function for as.data.frame()
```
for which definitions are given in the help file `?create_intervals`.

## MCMC samples for multiple years

We now present calculations for  multiple years
of MCMC samples of relative biomass.

To create ETIs and HDIs for estimates for each year:
```{r hake_mcmc2}
res_all_years <- create_intervals(petrale_mcmc,
                                  credibility = cred)
```
This gives a list object -- see the rendered `results.html` vignette for
details (such as the intervals for all years are convenienty saved in the single
tibble `res_all_years$intervals_all`.

Plot the full time series like in our manuscript figure:
```{r hake_mcmc7, echo = FALSE}
plot(res_all_years,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8))
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
```

Reproduce in the style of Figure 39 the Petrale Sole stock assessment (excluding 2024):
```{r plotlikeassessment, fig.height = 10, echo = FALSE}
to_plot <- res_all_years$intervals_all %>%
  filter(quantity != 2024)
par(mfrow = c(2,1))
plot(to_plot$quantity,
     to_plot$median,
     type = "o",
     pch = 16,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8),
     main = "Equal-tailed Intervals")
polygon(x = c(to_plot$quantity,
              rev(to_plot$quantity)),
        y = c(to_plot$eti_lower,
              rev(to_plot$eti_upper)),
        col = "lightblue",
        border = NA)
# Redo to be on top
points(to_plot$quantity,
       to_plot$median,
       type = "o",
       pch = 16)
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
add_minor_tickmarks(y_tick_by = 1)

plot(to_plot$quantity,
     to_plot$median,
     type = "o",
     pch = 16,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8),
     main = "Highest Density Intervals")
polygon(x = c(to_plot$quantity,
              rev(to_plot$quantity)),
        y = c(to_plot$hdi_lower,
              rev(to_plot$hdi_upper)),
        col = "lightblue",
        border = NA)
# Redo to be on top
points(to_plot$quantity,
       to_plot$median,
       type = "o",
       pch = 16)
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
add_minor_tickmarks(y_tick_by = 1)
```

Adapted from hake vignette: whether the relative spawning biomass is above or below the management reference points
of 0.4 and 0.8 is key in forming stakeholders' perception of the health of the hake
stock.

Years for which ETI is above 0.4 but HDI is below:
```{r whichyears}
dplyr::filter(res_all_years$intervals_all,
              eti_lower > 0.4,
              hdi_lower < 0.4) %>%
  select(quantity,
         eti_lower,
         hdi_lower)
```
Thus, the perception of stock status can depend upon the
definition of credible intervals (ETI versus HDI).


Only one year for which lower end of ETI is above 0.8 but HDI is below (the
stock is increasing fairly quickly around this time):
```{r whichyearsusr}
dplyr::filter(res_all_years$intervals_all,
              eti_lower > 0.8,
              hdi_lower < 0.8) %>%
  select(quantity,
         eti_lower,
         hdi_lower)
```

Can see from the time series that upper end of both intervals are always above
0.8.

From the time series figure, there looks to be quite a difference between
intervals in the 1960s and 1970s. Plotting, say, 1969:
```{r petrale_1969, fig.height = 8, echo = FALSE}
ind_1969 <- which(years == 1969)
x_lim <- c(0, 2)
y_lim = c(0, 5)
y_arrow = 4.3

par(mfrow = c(2, 1))
plot(res_all_years$res_all[[ind_1969]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.5)
     # main = "Equal-tailed Interval")
plot(res_all_years$res_all[[ind_1969]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.5)
     # main = "Highest Density Interval")
```

And also plot the same for 1979:
```{r petrale_1979, fig.height = 8, echo = FALSE}
ind_1979 <- which(years == 1979)
x_lim <- c(0, 2.5)
y_lim = c(0, 5)
y_arrow = 4.3

par(mfrow = c(2, 1))
plot(res_all_years$res_all[[ind_1979]],
     type = "eti",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.5)
     # main = "Equal-tailed Interval")
plot(res_all_years$res_all[[ind_1979]],
     type = "hdi",
     xlim = x_lim,
     ylim = y_lim,
     interval_arrows = TRUE,
     xlab = x_lab,
     y_arrow = y_arrow,
     arrowhead_gap = 0,
     x_minor_ticks_by = 0.25,
     y_minor_ticks_by = 0.5)
     # main = "Highest Density Interval")
```



To see the main values of interest, namely the ETIs and HDIs and the difference
between their widths:
```{r hake_mcmc5}
summary_table(res_all_years)
```

The sum of the differences over all years between the width of the ETI and width
of the HDI is:
```{r hake_mcmc6}
res_all_years$intervals_all %>%
  dplyr::pull(width_diff) %>%
  sum()
```
although relative biomass is probably not really a quantity that should be summed.

## 50% credible intervals

How about 50% credible intervals like I did quickly for hake (for which results
were very surpising):
```{r int50}
res_all_years_50 <- create_intervals(petrale_mcmc,
                                     credibility = 0.5)
```

```{r int50plot, echo = FALSE}
plot(res_all_years_50,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 4))
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
```

## 90% credible intervals

How about 90% credible intervals:
```{r int90}
res_all_years_90 <- create_intervals(petrale_mcmc,
                                     credibility = 0.9)
```

```{r int90plot, echo = FALSE}
plot(res_all_years_90,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8))
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
```

## Original 95% ETI compared to 90% HDI

How about comparing original 95% ETI, with new 90% HDI:
```{r plotlikeassessment2, fig.height = 10, echo = FALSE}
to_plot <- res_all_years$intervals_all %>%
  filter(quantity != 2024)
par(mfrow = c(2,1))
plot(to_plot$quantity,
     to_plot$median,
     type = "o",
     pch = 16,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8),
     main = "Original Equal-tailed Intervals, 95%")
polygon(x = c(to_plot$quantity,
              rev(to_plot$quantity)),
        y = c(to_plot$eti_lower,
              rev(to_plot$eti_upper)),
        col = "lightblue",
        border = NA)
# Redo to be on top
points(to_plot$quantity,
       to_plot$median,
       type = "o",
       pch = 16)
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
add_minor_tickmarks(y_tick_by = 1)

to_plot <- res_all_years_90$intervals_all %>%
  filter(quantity != 2024)
plot(to_plot$quantity,
     to_plot$median,
     type = "o",
     pch = 16,
     xlab = "Year",
     ylab = x_lab,
     ylim = c(0, 8),
     main = "Highest Density Intervals but 90%")
polygon(x = c(to_plot$quantity,
              rev(to_plot$quantity)),
        y = c(to_plot$hdi_lower,
              rev(to_plot$hdi_upper)),
        col = "lightblue",
        border = NA)
# Redo to be on top
points(to_plot$quantity,
       to_plot$median,
       type = "o",
       pch = 16)
abline(h = 0.4, col = "red", lty = 2)
abline(h = 0.8, col = "black", lty = 2)
add_minor_tickmarks(y_tick_by = 1)
```

```{r hake_mcmc8, eval = FALSE, echo = FALSE}
# To see all the ETIs and medians (useful for checking with any published tables)
res_all_years$intervals_all %>%
  dplyr::select(quantity, eti_lower, median, eti_upper) %>%
  a() %>%
  round(2)
```
