---
title: "Extra results and examples of code functionality"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extra results and examples of code functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, rendering, echo = FALSE, eval = FALSE}
rmarkdown::render("results-extra.Rmd")
# to build, or click the knit button in RStudio
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  cache_path = "results-extra-cache/",
  fig.path = "results-extra-figs-cache/",
  fig.width = 7,
  fig.height = 6
)
```

This vignette follows on from the
[results.html](http://htmlpreview.github.io/?https://github.com/andrew-edwards/hdiAnalysis/blob/main/vignettes/results.html)
vignette for our [hdiAnalysis R package](https://github.com/andrew-edwards/hdiAnalysis?tab=readme-ov-file#hdianalysis), including:

- plotting of all density functions when multiple years of data are analysed
 (specifically all years of recruitment shown in Figure 2A).
- details of plotting Fig. S.1, demonstrating of how ranges a and b (Fig. 1B)
  and ranges c and d are defined
- exploration of differences between ETIs and HDIs for hake relative biomass and
  recruitment when using 90%,
  80%, 66% and 50% credible intervals (95% was used for the main results).
- calculation on a longer MCMC chain for the hake assessment results.
- exporation of additional options for calculating HDIs, based on the kernel
density estimate of the MCMC sample.
- session information (documenting R and package version numbers).

Set up:
```{r setup}
library(hdiAnalysis)
library(dplyr)
```

## MCMC samples for multiple years -- recruitments as in Figure 2A but plotting the full distributions

It is instructive to look at the full distributions of recruitment for each
year.

For the multiple years of MCMC samples of hake recruitment, we calculate the
results as in `results.Rmd` (see that for more details) and plot the full
distributions with ETI and HDI for each year.
```{r hake_mcmc2}
res_all_years <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                -"Virgin"))
```

The default plot (which uses `plot.intervals_density_list()`) is that shown in Figure 2A:
```{r hake_mcmc7}
plot(res_all_years,
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
```

To plot all density plots and ETIs for each year (equivalent to Figure 1A), the `type` argument is used
(and we have changed the `fig.height` for the chunk, and specified `mfrow` as an
argument, and fixed `xlim` so that it is the same for each plot). The year is
automatically shown in each figure.
```{r hake_mcmc9, fig.height = 80}
plot(res_all_years,
     type = "eti",
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)",
     mfrow = c(30, 2))  # 59 default
```

For the equvalent figures for HDIs, just change the `type` argument:
```{r hake_mcmc9hdi, fig.height = 80}
plot(res_all_years,
     type = "hdi",
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)",
     mfrow = c(30, 2))  # 59 default
```

## Detailed explanation of calculation of ranges 'a', 'b', 'c', and 'd' in Figure S.1

The calculation of ranges 'a' and 'b' in Fig. 1A is described with respect to
the more detailed Fig. S.1A, which is calculated here:

```{r ab}
res_vec <- create_intervals(rec_2021)

plot(res_vec,
     type = "eti",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of fish)",
     arrowhead_gap = 0.1,
     explanatory_lines_a_b = TRUE)
```

This shows the ETI in Fig. 1A but with extra lines.
Two of the vertical grey lines show the
recruitment values at the ends of the ETI (at the red/blue boundaries), denoting
the right ends of 'a' and 'b'. Horizontal grey lines show the densities at these
two values; the intersection of these lines with the opposite limb of the
distribution gives the other two vertical grey lines, defining the left ends
of 'a' and 'b'. Therefore, 'a' and 'b' both span the same range of densities on
the y-axis,
but 'a' is outside of the ETI while 'b' is inside of it. (See also Supporting Information).

Another undesirable feature of ETIs (for asymmetric distributions) is that there
we can construct ranges `c` and `d` for which all values in `c` are more likely
than those in `d`, yet `c` is outside the ETI and `d` is inside it:
```{r green}
plot(res_vec,
     type = "eti",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of fish)",
     arrowhead_gap = 0.1,
     explanatory_lines_a_b = FALSE,
     show_a_b = FALSE,
     explanatory_lines_extra = TRUE)
```

Here, the horizontal green line is set (as an example) halfway between the
densities at the ends of the ETI (horizontal grey lines are as in the previous
figure). Its intersections with the density curve define the left ends of `c`
and `d`, and all values within `c` are more likely than all those in `d`, but
`c` is outside the ETI and `d` is inside it. The green line can be placed at any
point between the two grey lines to create such ranges `c` and `d`. This includes at
the density corresponding the HDI, as in Figure 12.2 of Kruschke (2015; Doing
Bayesian Data Analysis: A Tutorial with R, JAGS, and Stan. 2nd Edition).

The above plot settings are incorporated into our function `figure_s1()` for
reproducing Fig. S1.

## Exploration of alternative levels of confidence

Here we repeat Fig. 2 for hake recruitment and relative spawning biomass but for alternative levels of confidence to the 95% used in the hake
assessment. We repeat Fig. 2B using 90%, 80%, 66% and 50% credible intervals.

```{r relbiomassalt}
res_all_years_rec_95 <- res_all_years

res_all_years_rec_90 <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                       -"Virgin"),
                                         credibility = 0.90)

res_all_years_rec_80 <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                       -"Virgin"),
                                         credibility = 0.80)

res_all_years_rec_66 <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                       -"Virgin"),
                                         credibility = 0.66)

res_all_years_rec_50 <- create_intervals(dplyr::select(hake_recruitment_mcmc,
                                                       -"Virgin"),
                                         credibility = 0.50)

res_all_years_rel_biomass_95 <- create_intervals(hake_relative_biomass_mcmc,
                                                 credibility = 0.95)

res_all_years_rel_biomass_90 <- create_intervals(hake_relative_biomass_mcmc,
                                                 credibility = 0.90)

res_all_years_rel_biomass_80 <- create_intervals(hake_relative_biomass_mcmc,
                                                 credibility = 0.80)

res_all_years_rel_biomass_66 <- create_intervals(hake_relative_biomass_mcmc,
                                                 credibility = 0.66)

res_all_years_rel_biomass_50 <- create_intervals(hake_relative_biomass_mcmc,
                                                 credibility = 0.50)
```

The figures for each credibility in turn are (plotting code is hidden) for
recruitments, including the 95% plot from Fig. 2A):
```{r recruitmentcredsplot, echo = FALSE, fig.height = 20}
par(mfrow = c(5, 1))
plot(res_all_years_rec_95,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 95%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_90,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 90%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_80,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 80%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_66,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 66%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_50,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 50%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)
```

And for relative spawning biomass:
```{r relbiomasscredsplot, echo = FALSE, fig.height = 20}
par(mfrow = c(5, 1))
plot(res_all_years_rel_biomass_95,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 95%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_90,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 90%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_80,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 80%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_66,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 66%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_50,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 50%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)
```

```{r, reccredspdfs, echo = FALSE}
# Create pdf for recruitment plots, copying the above code to here:
pdf_height = 8
pdf_width = 114/25.4

pdf(file = "figure-recruitment-credibilities.pdf",
    height = pdf_height,
    width = pdf_width,
    paper="special")

par(mai = c(0.32, 0.35, 0.15, 0.12),
    mfrow = c(5,1),
    mgp = c(1.6,0.5,0),
    cex = 0.6)

plot(res_all_years_rec_95,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 95%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_90,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 90%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_80,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 80%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_66,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 66%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rec_50,
     ylim = c(0, 32),
     xlab = "Year",
     ylab = "Recruitment (billions of fish)")
mtext("Credibility = 50%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

dev.off()
```

```{r, relbiomasscredspdfs, echo = FALSE}
# Create pdf for relative biomass plots, copying the above code to here:
pdf(file = "figure-relative-biomass-credibilities.pdf",
    height = pdf_height,
    width = pdf_width,
    paper="special")

par(mai = c(0.32, 0.35, 0.15, 0.12),
    mfrow = c(5,1),
    mgp = c(1.6,0.5,0),
    cex = 0.6)

plot(res_all_years_rel_biomass_95,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 95%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_90,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 90%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_80,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 80%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_66,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 66%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

plot(res_all_years_rel_biomass_50,
     xlim = c(2010, 2027),
     ylim = c(0, 2.6),
     add_line_at_0.4 = TRUE,
     inc = 0.05,
     leg_loc = "topleft",
     xlab = "Year",
     ylab = "Relative spawning biomass")
mtext("Credibility = 50%",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)
dev.off()
```


## Longer chain of 32,000 MCMC samples for hake recruitment in 2021

We also analysed an experimental longer chain of 32,000 MCMC samples (that took
several days to run), and results are similar to the standard chain of 8,000
samples. Code is not run here (since the data object is large and so not included
in the package) but results are presented below.

```{r longchain, echo = FALSE, eval = FALSE}
vec_long <- dplyr::pull(hake_recruitment_mcmc_14_long,
                        "2021")

res_vec_long <- create_intervals(vec_long)

summary_table(res_vec_long)

plot(res_vec_long,
     type = "hdi",
     xlim = c(0, 40),
     ylim = c(0, 0.11),
     interval_arrows = TRUE,
     xlab = "Recruitment (billions of fish)",
     arrowhead_gap = 0.1)
# Density kernel is a bit smoother
```

| Quantity| Value | Interval width |
| :-------| -----:| -----:|
| Median  | 10.20 | -- |
| ETI     | 4.05 -- 29.53 | 25.49 |
| HDI     | 2.89 -- 25.23 | 22.34 |
| Range a | 2.27 -- 4.05 | 1.78 |
| Range b | 16.02 -- 29.53 | 13.52 |

The slight differences between the HDI for this chain and the one used in the
main results are small
compared to the difference to the ETI (for either chain). The ETIs change
slightly between chains due to randomness between them, and the median even
changes very slightly.

Also, the density values at the endpoints of the HDI are still not equal (they
should be theoretically):
```{r longchain2, eval = FALSE}
> res_vec_long$interval$y_hdi_lower
[1] 0.01026998
> res_vec_long$interval$y_hdi_upper
[1] 0.007201427
```
These are similar to the results for the original chain, suggesting long MCMC
chains will not result in the densities at the HDI endpoints being exactly
equal. This is likely due to the true bumpy shape of the posterior and the
approximations from the kernel density estimate required to estimate the densities.

Also, the integral over the HDI is estimated to be 94.7%, which is a slight
improvement to the 94.5% for the original chain of 8,000 samples, but still not
exactly 95%, such that the minor issue of the HDI not being exactly 95% is not
resolved by using the longer chain.

## Computing the HDI on the kernel density estimate

Given the minor issue in Fig 1B that the densities at the endpoints of the HDI
are not exactly equal, but they should be in theory, we also tried another
approach to computing the HDI. This involved first calculating the kernel
density estimate on the MCMC sample, and then using that to calculate the
HDI. The HDInterval package includes the function
\texttt{HDInterval::hdi.density()} to do this, which we incorporate into our
`create_intervals()` function with the argument `density = TRUE`.

So repeating our Fig. 1B of 2021 hake recruitment, but using the density option,
we obtain:

```{r density}
res_2021_density <- create_intervals(rec_2021,
                                     density = TRUE)
res_2021_density$intervals %>% a()
```
with the main results being:
```{r densitymain, results = "asis"}
summary_table(res_2021_density)
```
The HDI is shifted downwards slightly from our default calculation of
2.80--25.18, and is slightly narrower.

The equivalent plot to our Fig. 1B for the calculated HDI is
```{r densityplot}
plot(res_2021_density,
     xlim = c(0,40),
     xlab = "Recruitment (billions of fish)")
```
in which we can see that the heights of the two endpoints are equal.

Furthermore, the integral over the
HDI is the theoretical 95%:
```{r density4}
res_2021_density$intervals$integral_hdi
```

One technical issue we resolved was that the default of 512 equally spaced points (`n`) at which to compute the kernal density
estimate in the `density()` function, did not resolve the issue of unequal
densities at the endpoints of the HDI.
Setting the number of points to 512 here shows this:
```{r n512}
res_2021_n_512 <- create_intervals(rec_2021,
                                   n = 512)
plot(res_2021_n_512,
     xlim = c(0, 40),
     xlab = "Recruitment (billions of fish)")
```

We found that increasing `n` to 100,000 resolved this (this does not increase the wiggliness of the kernel density estimate,
just the number of points at which it is calculated), and hence is set as the
default in our `create_intervals()` function.

So while the density approach (with `n=100,000`) did seem preferable, given the equal densities at
the endpoints of the HDI, the following simple simulation experiment convinced
us to stick with the moving-window approach. We present these results for
interested users.

## Simulation experiment

We simulate 8,000 values from a lognormal distribution
with mean on the log scale of 4 and standard deviation of 0.5, somewhat
resembling our recruitment MCMC samples. We calculate the HDI using our default
approach and with the density approach just described. We repeated this
experiment 1,000 times, to compare the calculated HDI's with the known true
theoretical values (because we know the true underlying distribution, unlike for
our MCMC samples).

The true theoretical values are calculated from the
inverse cumulative distribution function for the lognormal
distribution, using the \texttt{HDInterval::hdi.function()} function.

```{r ints}
mean_log <- 4
sd_log <- 0.5
n <- 8000

# True theoretical values:
hdi_true <- HDInterval::hdi(qlnorm,
                            credMass = 0.95,
                            meanlog = mean_log,
                            sdlog = sd_log)
hdi_true

# Format into tibble output to match the $intervals output from `create_intervals()` to then
# easily append the results of the simulation:
ints_true <- tibble::tibble(median = qlnorm(0.5, mean_log, sd_log),
                            eti_lower = qlnorm(0.025, mean_log, sd_log),
                            eti_upper = qlnorm(0.975, mean_log, sd_log),
                            hdi_lower = hdi_true["lower"],
                            hdi_upper = hdi_true["upper"]) %>%
  dplyr::mutate(width_eti = eti_upper - eti_lower,
                width_hdi = hdi_upper - hdi_lower,
                width_diff = width_eti - width_hdi,
                a_lower = NA,
                b_lower = NA,
                i_eti_lower = NA,
                y_eti_lower = dlnorm(eti_lower,
                                     mean_log,
                                     sd_log),
                i_eti_upper = NA,
                y_eti_upper = dlnorm(eti_upper,
                                     mean_log,
                                     sd_log),
                i_hdi_lower = NA,
                y_hdi_lower = dlnorm(hdi_lower,
                                     mean_log,
                                     sd_log),
                i_hdi_upper = NA,
                y_hdi_upper = dlnorm(hdi_upper,
                                     mean_log,
                                     sd_log),
                hdi_height = NA,
                integral_full = NA,
                integral_eti = plnorm(eti_upper,
                                     mean_log,
                                     sd_log) -
                               plnorm(eti_lower,
                                      mean_log,
                                      sd_log),
                integral_hdi = plnorm(hdi_upper,
                                     mean_log,
                                     sd_log) -
                               plnorm(hdi_lower,
                                      mean_log,
                                      sd_log),
                warning = NA,
                allow_hdi_zero = NA)

ints_true %>% a()
```

We can see that the densities at the two ends of the HDI are (exactly) equal, as expected:
```{r densitiesequal}
ints_true$y_hdi_lower
ints_true$y_hdi_upper
```

Now to do 1,000 simulations, and compare the results to the theoretical values:
```{r simulations}
num_sims <- 1000
simulations <- list()
# Set up a tibble to then add results for each simulation as a new row, with the
# first row being the theoretical results obtained above:

# Results when not using density option (our default):
ints_simulations_density_false <- cbind(type = "theoretical",
                                      ints_true)

# Results when using density option:
ints_simulations_density_true <- ints_simulations_density_false # same format

for(i in 1:num_sims){
  set.seed(i)
  this_sim <- rlnorm(n,
                     mean_log,
                     sd_log)

  # Not using density:
  this_int_density_false <- create_intervals(this_sim,
                                           density = FALSE)

  ints_simulations_density_false <- rbind(ints_simulations_density_false,
                                        cbind(type = i,
                                              this_int_density_false$intervals))

  # Using density:
  this_int_density_true <- create_intervals(this_sim,
                                           density = TRUE)

  ints_simulations_density_true <- rbind(ints_simulations_density_true,
                                        cbind(type = i,
                                              this_int_density_true$intervals))
}

ints_simulations_density_false <- tibble::as_tibble(ints_simulations_density_false)
ints_simulations_density_true <- tibble::as_tibble(ints_simulations_density_true)

ints_simulations_density_false
ints_simulations_density_true
```

So we have 1,000 HDIs for each option (using `density` or not), so plot the
intervals and compare with the known values. Show the HDIs, including the first
one that represents the theoretical HDI which we illustrate with in red, and
plot vertical red lines for its endpoints, to compare with the simulated HDIs:
```{r plotints}
# Set up plot
plot(range(c(ints_simulations_density_false$hdi_lower,
             ints_simulations_density_false$hdi_upper)),
     c(1, nrow(ints_simulations_density_false)),
     type = "n",
     xlab = "95% interval",
     ylab = "Simulation row number",
     main = "Intervals with density = FALSE")

lines(c(ints_simulations_density_false[1, "hdi_lower"],
        ints_simulations_density_false[1, "hdi_upper"]),
      c(1, 1),
      col = "red")
for(i in 1:num_sims){
  lines(c(ints_simulations_density_false[i+1, "hdi_lower"],
        ints_simulations_density_false[i+1, "hdi_upper"]),
      c(i+1, i+1),
      col = "black")
}
abline(v = c(ints_simulations_density_false[1, "hdi_lower"],
             ints_simulations_density_false[1, "hdi_upper"]),
       col = "red")
```

These look okay, but to more clearly compare we show histograms of lower and
upper HDI values compared to true theoretical value (includes the
theoretical value in the histogram):
```{r checklower}
hist(ints_simulations_density_false$hdi_lower,
     xlab = "HDI lower value",
     main = "Lower HDI with density = FALSE")
abline(v = ints_simulations_density_false[1, "hdi_lower"], col = "red")

hist(ints_simulations_density_false$hdi_upper,
     xlab = "HDI upper value",
     main = "Upper HDI with density = FALSE")
abline(v = ints_simulations_density_false[1, "hdi_upper"], col = "red")
```

The actual numbers of simulated lower endpoints of HDI above the true value is
```{r calcnums}
sum(ints_simulations_density_false$hdi_lower[-1] > ints_simulations_density_false$hdi_lower[1])
```
and the number of upper endpoints above the true value is
```{r calcnumupper}
sum(ints_simulations_density_false$hdi_upper[-1] > ints_simulations_density_false$hdi_upper[1])
```
So not exactly half (`r num_sims/2`), but okay.

Now to do the same plots for the results when using the density approach to
calculating the HDIs:
```{r plotintsdens}
# Set up plot
plot(range(c(ints_simulations_density_true$hdi_lower,
             ints_simulations_density_true$hdi_upper)),
     c(1, nrow(ints_simulations_density_true)),
     type = "n",
     xlab = "95% interval",
     ylab = "Simulation row number",
     main = "Intervals with density = TRUE")

lines(c(ints_simulations_density_true[1, "hdi_lower"],
        ints_simulations_density_true[1, "hdi_upper"]),
      c(1, 1),
      col = "red")
for(i in 1:num_sims){
  lines(c(ints_simulations_density_true[i+1, "hdi_lower"],
        ints_simulations_density_true[i+1, "hdi_upper"]),
      c(i+1, i+1),
      col = "black")
}
abline(v = c(ints_simulations_density_true[1, "hdi_lower"],
             ints_simulations_density_true[1, "hdi_upper"]),
       col = "red")
```

These do not look great at the low end, as we can see from the histogram (the one
value on the right in the first panel is indeed the true theoretical value as
it's included in the tibble of lower HDI values):
```{r checklowerdens}
hist(ints_simulations_density_true$hdi_lower,
     xlab = "HDI lower value",
     main = "Lower HDI with density = TRUE")
abline(v = ints_simulations_density_true[1, "hdi_lower"], col = "red")

hist(ints_simulations_density_true$hdi_upper,
     xlab = "HDI upper value",
     main = "Upper HDI with density = TRUE")
abline(v = ints_simulations_density_true[1, "hdi_upper"], col = "red")
```

The actual numbers of simulated lower endpoints of HDI above the true value is
```{r calcnums2}
sum(ints_simulations_density_true$hdi_lower[-1] > ints_simulations_density_true$hdi_lower[1])
```
and the number of upper endpoints above the true value is
```{r calcnumupper2}
sum(ints_simulations_density_true$hdi_upper[-1] > ints_simulations_density_true$hdi_upper[1])
```

So these values are quite a way off being the desired half (`r
num_sims/2`). This convinced us to stick with the simpler no-density
moving-window method.

## Multimodal distributions

Multimodal distributions of MCMC output can occur from the underlying
stochasticity of MCMC results, and also from ensemble approaches whereby results
from multiple models are combined to provide model-averaged result. To
investigate the associated intervals, we first create a
simple bimodal distribution of simulated data and calculate an HDI using the moving-window approach:

```{r bimodal}
set.seed(42)
bimodal_sim <- c(rlnorm(n/2,
                        2.2,
                        sd_log/4),
                 rlnorm(n/2,
                        3.5,
                        sd_log/2))
bimodal_res <- create_intervals(bimodal_sim)
```

```{r bimodalres, results = "asis"}
summary_table(bimodal_res)
```

```{r bimodalfigs}
# Equal-tailed interval:
plot(bimodal_res,
     type = "eti",
     xlim = c(0, 70),
     ylim = c(0, 0.1),
     interval_arrows = TRUE,
     xlab = "Simulated data",
     y_arrow = 0.09,
     y_minor_ticks_by = 0.01,
     arrowhead_gap = 0.05,
     show_a_b = FALSE)

# Highest density interval using moving-window approach
plot(bimodal_res,
     type = "hdi",
     xlim = c(0, 70),
     ylim = c(0, 0.1),
     interval_arrows = TRUE,
     xlab = "Simulated data",
     y_arrow = 0.09,
     y_minor_ticks_by = 0.01,
     arrowhead_gap = 0.05)
```

As can be seen in this example, there is a region (around 17) within the HDI but
with a low density, and the difference between the densities at the endpoints of
the interval is quite different (though still less so than for the ETI). Thus, this is not strictly an HDI, but is still a valid credible
interval. This is using the default
moving-window approach. (Code to create the figure for the Supplementary
Material is next in this .Rmd but not shown in the rendered .html file).

```{r, figurebimodal, echo = FALSE}
# Create pdf for bimodal distribution
pdf_height = 4
pdf_width = 114/25.4

pdf(file = "figure-bimodal.pdf",
    height = pdf_height,
    width = pdf_width,
    paper="special")

par(mai = c(0.32, 0.3, 0.15, 0.12),
    mfrow = c(2,1),
    mgp = c(1.6,0.5,0),
    cex = 0.6)

# ETI:
plot(bimodal_res,
     type = "eti",
     xlim = c(0, 70),
     ylim = c(0, 0.1),
     interval_arrows = TRUE,
     xlab = "Simulated data",
     y_arrow = 0.09,
     y_minor_ticks_by = 0.01,
     arrowhead_gap = 0.05,
     show_a_b = FALSE)
mtext("A. Equal-tailed interval",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

# HDI using moving-window approach
plot(bimodal_res,
     type = "hdi",
     xlim = c(0, 70),
     ylim = c(0, 0.1),
     interval_arrows = TRUE,
     xlab = "Simulated data",
     y_arrow = 0.09,
     y_minor_ticks_by = 0.01,
     arrowhead_gap = 0.05)

mtext("B. Highest density interval",
      side = 3,
      adj = 0,
      cex = 0.7,
      line = 0.3)

dev.off()
```


To obtain disjoint intervals (and therefore highest density regions, not intervals) requires
using the density approach. This can be done using the original
`HDInterval::hdi()` function (because our `create_intervals()` is set up to only
deal with single intervals),
and allowing split (disjoint) intervals:
```{r region}
bimodal_region <- HDInterval::hdi(bimodal_res$density,  # Use already-calculated density
                                  allowSplit = TRUE)
bimodal_region
```
The results show two intervals that make up the highest density region.
We can manually create a figure, showing the highest density region in blue:
```{r regionfig}
dens <- bimodal_res$density

plot(bimodal_res,
     type = "eti",
     col_tail = "blue",
     xlim = c(0, 70),
     ylim = c(0, 0.09),
     interval_arrows = FALSE,
     xlab = "Simulated data",
     y_minor_ticks_by = 0.01)

# Lower left tail
polygon(c(dens$x[dens$x <= bimodal_region[1, "begin"]],
          bimodal_region[1, "begin"],
          bimodal_region[1, "begin"]),
        c(dens$y[dens$x <= bimodal_region[1, "begin"]],
          attr(bimodal_region, "height"),
          0),
        col = "red",
        border = "red",
        lwd = 0.4,
        main = "")
# Middle interval that is outside of highest density region
polygon(c(dens$x[dens$x >= bimodal_region[1, "end"] & dens$x <= bimodal_region[2, "begin"]],
          bimodal_region[2, "begin"],
          bimodal_region[2, "begin"],
          bimodal_region[1, "end"]),
        c(dens$y[dens$x >= bimodal_region[1, "end"] & dens$x <= bimodal_region[2, "begin"]],
          attr(bimodal_region, "height"),
          0,
          0),
        col = "red",
        border = "red",
        lwd = 0.4,
        main = "")

# Upper right left tail
polygon(c(bimodal_region[2, "end"],
          dens$x[dens$x >= bimodal_region[2, "end"]],
          bimodal_region[2, "end"]),
        c(attr(bimodal_region, "height"),
          dens$y[dens$x >= bimodal_region[2, "end"]],
          0),
        col = "red",
        border = "red",
        lwd = 0.4,
        main = "")
```

Thus, there are two disjoint intervals that make up the highest density region.
For stock assessments, it is unlikely
to be helpful to report such disjoint intervals for giving advice, and so we would
not recommend using them. Furthermore, in the simulations above we found the
density approach to not be accurate (for the sample size we tested). Therefore we
recommend investigating the use of HDIs or ETIs, noting that such HDIs are
still valid credible intervals and will be shorter than the corresponding ETIs.

## Session information

```{r sessioninfo}
sessionInfo()
```
