---
title: "Results for Pacific Herring, Petrale Sole, and Pacific Cod"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Results for Pacific Herring, Petrale Sole, and Pacific Cod}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, rendering, echo = FALSE, eval = FALSE}
rmarkdown::render("results-herring-petrale-cod.Rmd")
# to build, or click the knit button in RStudio
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  cache_path = "results-herring-petrale-cod-cache/",
  fig.path = "results-herring-petrale-cod-figs-cache/",
  fig.width = 7,
  fig.height = 6
)
```

Set up:
```{r setup}
library(hdiAnalysis)
```

We generate plots for three further stocks here, as documented in the
Supplementary Material. We keep this code short -- for further investigations
see the code in the respective species-specific code in `report/` or use further
code in the other vignettes.

## Pacific Herring

Using results for the Strait of Georgia stock from Fisheries and Ocean Canada's
[Stock status update with application of management procedures for Pacific Herring (*Clupea pallasii*) in British Columbia: status in 2023 and
forecast for 2024](https://publications.gc.ca/collections/collection_2024/mpo-dfo/fs70-7/Fs70-7-2024-001-eng.pdf).

HERE - use herring results as tempalte, simplifying down like below.

## Pacific Cod

Using results for Area 3CD from Fisheries and Ocean Canada's [Status update of
Pacific Cod (*Gadus Macrocephalus*) off the west coast of Vancouver Island in
2023](https://publications.gc.ca/collections/collection_2024/mpo-dfo/fs70-7/Fs70-7-2024-003-eng.pdf),
with MCMC results provided by Robyn Forrest.

### Recruitment

To create ETIs and HDIs for estimates for each year:
```{r, recruitment}
credibility_cod <- 0.95
res_all_years_cod_rec <- create_intervals(cod_rec_mcmc,
                                          credibility = credibility_cod)
```

Full time series of recruitments using both types of intervals:
```{r, plotrec}
plot(res_all_years_cod_rec,
     xlab = "Year",
     ylab = "Recruitment (millions of fish)",
     ylim = c(0, 160),
     y_tick_by = 10,
     add_line = 0,
     add_line_col = "grey",
     add_line_lty = 1)
```

Some pulling down of HDIs compared to ETIs, particularly 2009, 2013, 2022 and earlier
years. To see the density plot for each year (not run here):

```{r codallhdi, fig.height = 88, eval = FALSE}
plot(res_all_years_cod_rec,
     type = "hdi",
     xlim = c(0, 120),
     xlab = "Recruitment (millions of fish)",
     mfrow = c(33, 2))  # 65 default
```

The sum of the differences over all years between the width of the ETI and width
of the HDI is (in millions of fish)
```{r coddiffs}
res_all_years_cod_rec$intervals_all %>%
  dplyr::pull(width_diff) %>%
  sum()
```
so smaller than for hake, as expected due to the smaller stock size.

### Spawning biomass relative to reference points

Relative to the upper stock reference point, above which the stock is considered
to be healthy, the intervals of relative biomass for all years are calculated with:
```{r calcusr}
res_all_years_rel_usr <- create_intervals(cod_biomass_over_usr,
                                          credibility = credibility_cod)
```

(The Warnings are due to not being able to compute lower end of interval a).

Plot the full time series like in our manuscript figure:
```{r, plotusr}
plot(res_all_years_rel_usr,
     xlab = "Year",
     ylab = "Spawning biomass relative to USR",
     ylim = c(0, 3.25),
     y_tick_by = 0.25,
     add_line = 1,
     add_line_col = "green")
```
So there are only relatively minor changes, with only one year for which the
interval overlapping or not overlapping the critical value of 1 changes with the
definition of interval:
```{r codusroverlap}
dplyr::filter(res_all_years_rel_usr$intervals_all,
              eti_lower > 1,
              hdi_lower < 1)
dplyr::filter(res_all_years_rel_usr$intervals_all,
              eti_upper > 1,
              hdi_upper < 1)
```
So in the case the perception of stock status does not really change when using
HDI instead of ETI, presumably because the distributions look somewhat symmetric
(at least the median is roughly in the middle of the intervals).

Relative to the lower reference point, below which the stock is considered
to be in a critical zone, the intervals of relative biomass for all years are
calculated here. The lower reference point for this stock is defined as the
biomass in 1986 (the lowest estimated biomass agreed upon as
an undesirable state to be avoided), and so this year is removed from the
interval calculations because the relative biomass values would
be 1 for each MCMC sample in 1986 (since the biomass would be scaled by
itself, by definition the stock is at the reference point in 1986):
```{r, lrp}
res_all_years_rel_lrp <- create_intervals(dplyr::select(cod_biomass_over_lrp,
                                                        -"1986"),
                                          credibility = credibility_cod)
```
(Warnings are same as earlier.)

Plot the full time series like in our manuscript figure:
```{r, plotlpr, echo = FALSE}
plot(res_all_years_rel_lrp,
     xlab = "Year",
     ylab = "Spawning biomass relative to LRP",
     ylim = c(0, 12),
     y_tick_by = 0.5,
     add_line = 1,
     add_line_col = "red")
```
Again, the intervals do not change greatly between the ETIs and HDIs due to
apparent symmetry.

## Absolute spawning biomass

Given the symmetry noted above for relative biomass, it is slightly surprising
that the intervals do change somewhat when looking at absolute spawning biomass:

```{r resallyearsabsbiomass}
res_all_years_cod_biomass <- create_intervals(cod_biomass_mcmc,
                                              credibility = credibility_cod)
plot(res_all_years_cod_biomass,
     xlab = "Year",
     ylab = "Spawning biomass (1000 t)",
     ylim = c(0, 200),
     y_tick_by = 10)
```


## Session information

```{r sessioninfo}
sessionInfo()
```
